Building Sustainable Software - The CSDMS Approach

### 2014-11-15 hutton_et_al_2014_wssspe.tex
Updated with reviewer comments.
Rewrote the Education section and changed a few items in the WTM
section. Most importantly, added Scott as author!

\author{
  Eric W. H. Hutton$^{\dag}$,
  Mark D. Piper$^{\dag}$,
  Irina Overeem$^{\dag}$,
  Albert J. Kettner$^{\dag}$,
  James P. M. Syvitski$^{\dag}$,
}

@@

\item \emph{Accessibility}. As a web-based application, if you have access
      to the Internet, you have access to \wmt{}.

@@

\item \emph{Portability}. \wmt{} has a native JavaScript interface, so it can
      be accessed on any modern web browser

\item \emph{Maintenance}. Because modern browsers tend to adhere to web
      standards, which lead to fewer cross-compatibility issues than
      operating systems, only one version of WMT needs to be developed
      and maintained.

@@

\section{Education and Evaluation of \csdms{} Tools}
\label{sec:education}

%\wmt{} provides the ability to run models with an easy to use graphical
%interface and intends to serve novice and more experienced modelers alike.
%The WMT beta version was released in May 2014  as a sequel to the CMT.

\csdms{} constantly evaluates the effectiveness of its tools and standards
through user feedback from clinics and tutorials and one-on-one interviews
with \csdms{} members.  Modeling courses at the University of Colorado,
clinics at the International Summer Institute at the University of Minnesota,
and clinics at \csdms{} Annual Meetings are the dominant forums for
this testing and feedback.

Over the last 4 years over 150 graduate students - from geography, hydrology,
geology, oceanography, earth sciences and engineering departments - have
enrolled in intensive \csdms{}-run modeling courses. These courses
provide an excellent use case for us to test usability and access to our
tools. Students reported in post-course surveys that they were unfamiliar with
the use of models on a high performance computing cluster beforehand (rated
their familiarity as 1.5 average on a Likert scale of 1-5), but ended the
courses with familiarity and being more comfortable with running and changing
simulations as well as visualizing results (average rating approaches 4)
~\cite{overeem2013strategies}.

Course and clinic materials are posted online as educational resources.
\csdms{} develops \wmt{}-specific help pages for all models currently
available as components, so that users can easily learn about model
parameters and input files. In addition, we are developing web-based video
tutorials that show how to run models.

>>>>

\author{
  Eric W. H. Hutton$^{\dag}$,
  Mark D. Piper$^{\dag}$,
  Scott D. Peckham$^{\dag}$,
  Irina Overeem$^{\dag}$,
  Albert J. Kettner$^{\dag}$,
  James P. M. Syvitski$^{\dag}$,
}

@@

\item \emph{Accessibility}. As a web-based application, users with internet
      access, have access to \wmt{}.

@@

\item \emph{Portability}. \wmt{} has a native JavaScript interface.
      JavaScript is an open standard and almost every computer is equipped
      with a JavaScript-capable web browser.

%\item \emph{Maintenance}. Because modern browsers tend to adhere to web
%      standards, which lead to fewer cross-compatibility issues than
%      operating systems, only one version of WMT needs to be developed
%      and maintained.

@@

\section{Education}
\label{sec:education}

%\csdms{} constantly evaluates the effectiveness of its tools and standards
%through user feedback from clinics and tutorials and one-on-one interviews
%with \csdms{} members.  Modeling courses at the University of Colorado,
%clinics at the International Summer Institute at the University of Minnesota,
%and clinics at \csdms{} Annual Meetings are the dominant forums for
%this testing and feedback.

It has been our experience that new graduate students in the Earth sciences
begin with little or no programming experience. However, their interest in
programming is high and they see the importance of programming with regard
to their thesis topics. More often than not, programming courses are not
offered as part of their coursework and so, out of necessity to solve a
particular problem, students end up teaching themselves how to program.

We believe providing these students with even a small base in programming and
guidelines on best practices will go a long way. We have hosted several
programming workshops, including one through Software Carpentry. The workshops
were such a success that we are now planning an annual Software Carpentry
bootcamp that is specifically targeted to incoming graduate students. The
workshops fill up nearly immediately, which reflects a strong interest and
need for basic software skills. We believe these workshops strike a good
balance between teaching students about computer science topics and
maintaining a pragmatic focus on problem solving.

%Over the last 4 years over 150 graduate students - from geography, hydrology,
%geology, oceanography, earth sciences and engineering departments - have
%enrolled in intensive \csdms{}-run modeling courses. These courses
%provide an excellent use case for us to test usability and access to our
%tools. Students reported in post-course surveys that they were unfamiliar with
%the use of models on a high performance computing cluster beforehand (rated
%their familiarity as 1.5 average on a Likert scale of 1-5), but ended the
%courses with familiarity and being more comfortable with running and changing
%simulations as well as visualizing results (average rating approaches 4)
%~\cite{overeem2013strategies}.

%Course and clinic materials are posted online as educational resources.
%\csdms{} develops \wmt{}-specific help pages for all models currently
%available as components, so that users can easily learn about model
%parameters and input files. In addition, we are developing web-based video
%tutorials that show how to run models.
